---
title: "Create 2022 Data - Crime"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook creates the crime data for the Spring 2022 final projects.

## Dependencies
This notebook requires:

```{r load-packages, messages = FALSE}
# tidystl packages
library(compstatr)

# tidyverse packages
library(dplyr)
library(readr)

# other packages
library(here)
library(testthat)
```

## Load Raw Data
First, we load year-list objects using `cs_load_year()`. The raw data have been manually downloaded from SLMPD's open data portal.

```{r load-data}
data2019_raw <- cs_load_year(here("data", "raw", "2019"))
data2018_raw <- cs_load_year(here("data", "raw", "2018"))
data2017_raw <- cs_load_year(here("data", "raw", "2017"))
data2016_raw <- cs_load_year(here("data", "raw", "2016"))
data2015_raw <- cs_load_year(here("data", "raw", "2015"))
```

## Prep Data
Next, we'll validate and prepare our data.

### 2019
We validate the data to make sure it can be collapsed using `cs_validate_year()`:

```{r validate-data19}
cs_validate(data2019_raw, year = "2019")
```

Since the validation result is a value of `TRUE`, we can proceed to collapsing the year-list object into a single tibble with `cs_collapse()` and then stripping out crimes reported in 2018 for earlier years using `cs_combine()`. We also strip out unfounded crimes that remain using `cs_filter_count()`:

```{r collapse-data19}
# collapse into single object
data2019_raw <- cs_collapse(data2019_raw)

# combine and filter
cs_combine(type = "year", date = 2019, data2019_raw) %>%
  cs_filter_count(var = count) -> data2019
```

The `data2019` object now contains only crimes reported in 2019.

### 2018
We'll repeat the validation process with the 2018 data:

```{r validate-data18}
cs_validate(data2018_raw, year = "2018")
```

Since the validation result is a value of `TRUE`, we can proceed to collapsing the year-list object into a single tibble with `cs_collapse()` and then stripping out crimes reported in 2018 for earlier years using `cs_combine()`. We also strip out unfounded crimes that remain using `cs_filter_count()`:

```{r collapse-data18}
# collapse into single object
data2018_raw <- cs_collapse(data2018_raw)

# combine and filter
cs_combine(type = "year", date = 2018, data2019_raw, data2018_raw) %>%
  cs_filter_count(var = count) -> data2018
```

The `data2018` object now contains only crimes reported in 2018.

### 2017
We'll repeat the validation process with the 2017 data:

```{r validate-data17}
cs_validate(data2017_raw, year = "2017")
```

Since we fail the validation, we can use the `verbose = TRUE` option to get a summary of where validation issues are occurring. 

```{r validate-data17-verbose}
cs_validate(data2017_raw, year = "2017", verbose = TRUE)
```

The data for May 2017 do not pass the validation checks. We can extract this month and confirm that there are too many columns in the May 2017 release. Once we have that confirmed, we can standardize that month and re-run our validation.

```{r fix-may17}
# extract data and unit test column numbers
expect_equal(ncol(cs_extract_month(data2017_raw, month = "May")), 26)

# standardize months
data2017_raw <- cs_standardize(data2017_raw, month = "May", config = 26)

# validate data
cs_validate(data2017_raw, year = "2017")
```

We now get a `TRUE` value for `cs_validate_year()` and can move on to collapsing the 2017 and 2018 raw data objects to create a new object, `data2017`, that contains all known 2017 crimes including those that were reported or upgraded in 2018.

```{r collapse-data17}
# collapse into single object
data2017_raw <- cs_collapse(data2017_raw)

# combine and filter
cs_combine(type = "year", date = 2017, data2019_raw, data2018_raw, data2017_raw) %>%
  cs_filter_count(var = count) -> data2017
```

The `data2017` object now contains only crimes reported in 2017.

### 2016
We'll repeat the validation process with the 2016 data:

```{r validate-data16}
cs_validate(data2016_raw, year = "2016")
```

Since the validation process passes, we can immediately move on to creating our 2016 data object:

```{r collapse-data16}
# collapse into single object
data2016_raw <- cs_collapse(data2016_raw)

# combine and filter
cs_combine(type = "year", date = 2016, data2019_raw, data2018_raw, data2017_raw, 
           data2016_raw) %>%
  cs_filter_count(var = count) -> data2016
```

The `data2016` object now contains only crimes reported in 2016.

### 2015
We'll repeat the validation process with the 2015 data:

```{r validate-data15}
cs_validate(data2015_raw, year = "2015")
```

Since the validation process passes, we can immediately move on to creating our 2015 data object:

```{r collapse-data15}
# collapse into single object
data2015_raw <- cs_collapse(data2015_raw)

# combine and filter
cs_combine(type = "year", date = 2015, data2019_raw, data2018_raw, data2017_raw, 
           data2016_raw, data2015_raw) %>%
  cs_filter_count(var = count) -> data2015
```

The `data2015` object now contains only crimes reported in 2015.

## Clean-up Enviornment
We can remove the `_raw` objects at this point:

```{r remove-raw}
rm(data2015_raw, data2016_raw, data2017_raw, data2018_raw, data2019_raw)
```

## Create Single Table
Next, we'll create a single table of our data. We also subset columns to reduce the footprint of the table.

```{r collapse-all}
bind_rows(data2015, data2016, data2017, data2018, data2019) %>%
  select(date_occur, crime, description, ileads_address, ileads_street, x_coord, y_coord) -> crimes

write_csv(crimes, here("data", "clean", "crimes.csv"))
```
